---
title: "Ch3 Vectors"
output: 
  html_document:
    toc: true
---

<style type="text/css">
  body{
  font-size: 14pt;
  font-family: Georgia;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE)
```

## 3.1 Introduction

## 3.2 Atomic Vectors

### 3.2.5 Q1

```{r}
# To create a raw vector
charToRaw("I have a cat")
# To create a complex vector
complex(length.out = 3, real = 1, imaginary = 1)
```

### 3.2.5 Q2

```{r eval=FALSE}
c(1, FALSE)     # 1,  0
c("a", 1)       # "a", "1"
c(TRUE, 1L)     # 1L, 1L
```

### 3.2.5 Q3

```{r eval=FALSE}
1 == "1"        # True
-1 < FALSE      # True
"one" < 2       # False
```

### 3.2.5 Q4

As we can see, there is no coercion between NA and FALSE
```{r}
is.logical(NA)
c(FALSE, NA, NA_character_)   
```

### 3.2.5 Q5

- is.atomic() tests if an object is an atomic vector (as defined in Advanced R) or is NULL (!).
- is.numeric() tests if an object has type integer or double and is not of class factor, Date, POSIXt or difftime.
- is.vector() tests if an object is a vector (as defined in Advanced R) or an expression and has no attributes, apart from names.

```{r}
is.atomic(NULL)
x <- structure(1:3, start = 1, end = 0)
is.vector(x)
y <- matrix(1:9, nrow = 3, ncol = 3)
is.vector(y)
rm(x, y)
```

## 3.3 Attributes

### 3.3.4 Q1

```{r, eval=FALSE}
setNames = function (object = nm, nm) {
    names(object) <- nm
    object
}
setNames(1:3, c("a", "b", "c"))

unname = function (obj, force = FALSE) {
    if (!is.null(names(obj))) 
        names(obj) <- NULL
    if (!is.null(dimnames(obj)) && (force || !is.data.frame(obj))) 
        dimnames(obj) <- NULL
    obj
}
```

```{r}
# Compare names() and dimnames()
names(mtcars)
dimnames(mtcars)
```


### 3.3.4 Q2

```{r}
x <- 1:10
dim(x)
nrow(x)
ncol(x)
NROW(x)
NCOL(x)
```

### 3.3.4 Q4

```{r}
structure(1:5, comment1 = "my attribute")
x <- structure(1:5, comment = "my attribute")
x
comment(x)
attr(x, which = "comment")
attributes(x)
```

```{r, echo=FALSE}
rm(x)
```

## 3.4 S3 atomic vectors

### 3.4.5 Q1
```{r}
ch_vec <- c("a", "b", "a", "c", "b")
fac_vec <- factor(x = ch_vec, levels = c("a", "b", "c", "x"))
fac_vec
table(fac_vec)
table(fac_vec) |> typeof()          # integer vector
table(fac_vec) |> attributes()
```

```{r, echo=FALSE}
rm(ch_vec, fac_vec)
```

### 3.4.5 Q2

```{r}
f1 <- factor(letters)
f1
levels(f1) <- rev(levels(f1))
f1
```

### 3.4.5 Q3

```{r}
f2 <- rev(factor(letters))
f2
as.integer(f2)
f3 <- factor(letters, levels = rev(letters))
f3
as.integer(f3)
```

## 3.5 Lists

### 3.5.4 Q2

```{r}
l1 <- list("a", "b", "c")
as.vector(l1)               # as l1 is already a list, just not an atomic vector
unlist(l1)
```

### 3.5.4 Q3

```{r}
date <- as.Date("1970-01-02")
dttm_ct <- as.POSIXct("1970-01-01 01:00", tz = "UTC")
date
dttm_ct
c(date, dttm_ct)          # integer vector
c(dttm_ct, date)          # integer vector
```

```{r}
unclass(c(date, dttm_ct))
unlist(list(date, dttm_ct))
```

c() coerces types and strips time zones. Errors may have occurred in older R versions because of inappropriate method dispatch/immature methods. unlist() strips attributes.

```{r}
rm(date, dttm_ct, l1, f1, f2, f3)
```


## 3.6 Data frames and tibbles

```{r}
pacman::p_load(tibble)
df1 <- data.frame(
  x = letters[1:5],
  y = 1:5,
  z = (1:5) ^ 2
)
df1
is_tibble(df1)
attributes(df1)
attributes(df1)$class <-  c("data.frame", "tbl", "tbl_df")
df1
is_tibble(df1)
```

